Report on Late Winter2013 Production: Image Differencing

### 2013-03-26 W13report.tex
Wrote up section 5 on computational performance.

Section \ref{sec-analyticfp} discusses the theoretical foundation 
that sets the floor on the number of false positives we expect from a 
pure Gaussian random field.  The number of false positives is a strong
function of the seeing and and the confidence level of the detection.  
By comparing the emperical false positive rate to the analytic predictions
we can quantify how close we are getting to the theoretical floor in these
data.  Because the number of false positives is such a strong function of
detection threshold, misestimation of the signal to noise of the detections
has a large impact on how many false positives are observed.  The Figures
\ref{fig-fpthresh1}, \ref{fig-fpthresh2}, and \ref{fig-fpthresh3} show the 
effect of misestimating the noise in the detections for each of three different
seeing values (0.6\", 0.88\", 1.2\" respectively).

@@

For completeness, we note the median differences in the image plane at
the time of detection are 3--4 counts (or 0.2 sigma relative to the
empirical variance) for v6866601, 0.2 counts (0.02 sigma) for
v6866602, and less than 0.1 counts (sigma) for v6866603.

@@

Uh...

@@

\subsection{Astrometric Registration}
The original plan for this production was to use single deep exposures as 
templates.  This required that we treat the single exposures as coadds.  As 
such we warped the exposures to coadd patches for use in the image difference
task.  The register task is unable to solve a differential WCS if one of the images
has the CRPIX off the image.  This is the case for coadd patches.  To get around this
we resolved the coadd patches using a TAN-SIP solution and then used the register
task to solve for the differential WCS.  This worked fairly will part of the time 
yeilding residuals in the 40-100 mas range.  The rest of the time the 
residual was more like 250 mas range.  This is an order of magnitude larger than expected.

>>>>

Section \ref{sec-analyticfp} discusses the theoretical foundation 
that sets the floor on the number of false positives we expect from a 
pure Gaussian random field.  The number of false positives is a strong
function of the seeing and and the confidence level of the detection.  
By comparing the empirical false positive rate to the analytic predictions
we can quantify how close we are getting to the theoretical floor in these
data.  Because the number of false positives is such a strong function of
detection threshold, misestimation of the signal to noise of the detections
has a large impact on how many false positives are observed.  The Figures
\ref{fig-fpthresh1}, \ref{fig-fpthresh2}, and \ref{fig-fpthresh3} show the 
effect of misestimating the noise in the detections for each of three different
seeing values (0.6\", 0.88\", 1.2\" respectively).

@@

For completeness, we note the median differences in the image plane at
the time of detection are 3--4 counts (or 0.2 sigma relative to the
empirical variance) for v6866601, 0.2 counts (0.02 sigma) for
v6866602, and less than 0.1 counts (0.01 sigma) for v6866603.

@@

For this production, the understanding of the algorithm and
optimization of the false detection rates took priority over
optimizing the computational performance of the algorithms.  In
particular, debugging hooks were impolemented and workarounds of
problems were devised with the focus on understanding the core
Psf--matching algorithm.  Therefore there are several key ways in
which the production as--implemented will differ from how the code is
used in Nightly and Data Release Processing.  We outline these
differences below.

The first difference is that due to problems with the {\tt
  meas\_astrom} package, detailed in Section~\ref{subsec-astrom}, we
were unable to exercise the envisioned data flow where we query a
coadd repository for the template image.  Instead we used a single
deep {\tt calexp} as the template image.  The astrometric registration
of these images was trivial, as the images almost exactly overlap and
are exactly the same size.  Thus the timings of the {\tt
  imageDifference:register} task are {\it underestimates} of the
astrometric registration subtask computational performance.

Second, the core performance of the {\tt imageDifference:subtract}
task will be significantly impacted by debugging hooks that were put
in place specifically to generate the metrics being evaluated in this
production.  The core functionality, {\tt matchMaskedImages}, does the
convolution kernel fitting and actual convolution of the template
image, and will be least impacted by these add--ons.  The reported
timings of {\tt matchMaskedImages} are thus an {\it overestimate} of
the method's performance.

The reported {\tt imageDifference:detection} task performance is an
accurate representation of its computational requirements.  However,
the {\tt imageDifference:dipolemeasurement} task represents a
significant overestimate of the true dipole measurement requirements.
In particular, the Psf--flux dipole measurement is explicitly
sub--optimal, in that its centroid fitting algorithm is a 4--deep {\tt
  for} loop.

With these caveats, the timings of the {\tt imageDifference} subtasks
are presented in Table~\ref{tab-timings}.

@@

\subsection{Astrometric Registration \label{subsec-astrom}}
The original plan for this production was to use single deep exposures as 
templates.  This required that we treat the single exposures as coadds.  As 
such we warped the exposures to coadd patches for use in the image difference
task.  The register task is unable to solve a differential WCS if one of the images
has the CRPIX off the image.  This is the case for coadd patches.  To get around this
we resolved the coadd patches using a TAN-SIP solution and then used the register
task to solve for the differential WCS.  This worked fairly will part of the time 
yielding residuals in the 40-100 mas range.  The rest of the time the 
residual was more like 250 mas range.  This is an order of magnitude larger than expected.
